Distortion-corrected kernel density estimate on Riemannian manifolds

### 2022-09-26 paper/paper.Rmd
Upload twin peaks plots

```{r metadensity, fig.align = 'center', out.width = "80%", fig.cap="True density of the Gaussian mixture model of four kenels with means $(0.25, 0.25), (0.25, 0.75), (0.75, 0.25), (0.75, 0.75)$ and the same variance-covariance matrix $diag(0.02, 0.02)$. The colors indicate the density of the data and lower density points in darker colors are scattered both in the outer and center areas. The shapes indicate the four kernels.", echo = FALSE}
knitr::include_graphics("figures/truedensity_4kernels.png")
```

@@

```{r tpoutliers, fig.align = 'center', out.width = "95%", fig.cap="Highest density region plots of four manifold learning embeddings of the twin peak data. Variable KDE preforms better in finding kernel structures with ISOMAP and LLE, and in locating outliers with t-SNE and UMAP.", echo = FALSE}
knitr::include_graphics("figures/Twin Peak5levels_outliers_comparison_4ml_3cases_riem0_1.png")
```

@@

```{r tpcors, echo=FALSE, message=FALSE, eval=T}
load("figures/CorrelationTable_Twin Peak_4ml_riem0_1.rda")
col.max <- apply(cors, 2, max)
max.match <- rep(col.max, each = 2) == cors
data.frame(cors) %>% 
  mutate_if(is.numeric, format, digits = 3, nsmall = 3) %>% 
  kableExtra::kbl(caption = "Correlation between true density ranking and estimated density ranking for different manifold learning embeddings of the twin peak data. Variable bandwidth KDE outperfoms for LLE and UMAP, and LLE gives the highest rank correlation.", booktabs = TRUE, digits = 3, escape = FALSE) %>%
  # kable_styling(latex_options = "scale_down") %>%
  kable_paper(full_width = FALSE) %>%
  # column_spec(1, width = "6cm") %>% 
  column_spec(2, bold = max.match[,1]) %>%
  column_spec(3, bold = max.match[,2]) %>%
  column_spec(4, bold = max.match[,3]) %>%
  column_spec(5, bold = max.match[,4]) %>%
  column_spec(6, bold = max.match[,5])
```

@@

```{r tpisomapden, fig.align = 'center', out.width = "95%", fig.cap="Scatterplot of true density and estimated density ranks of ISOMAP embedding for KDE with both variable and fixed bandwidth. The four colors and shapes represents the four gaussian kernels in the 2-D meta data. Variable bandwidth shows a strong linear positive relationship.", echo = FALSE}
knitr::include_graphics("figures/Twin Peak_density_comparison_isomap_riem0_1_rank.png")
```

@@

>>>>

```{r metadensity, fig.align = 'center', out.width = "80%", fig.cap="True density of the Gaussian mixture model of four kenels with means $(0.25, 0.25), (0.25, 0.75), (0.75, 0.25), (0.75, 0.75)$ and the same variance-covariance matrix $diag(0.02, 0.02)$. The colors indicate the density of the data and lower density points in darker colors are scattered both in the outer and center areas. The shapes indicate the four kernels.", echo = FALSE}
knitr::include_graphics("figures/truedensity_twinpeaks_dc.png")
```

@@

```{r tpoutliers, fig.align = 'center', out.width = "95%", fig.cap="Highest density region plots of four manifold learning embeddings of the twin peak data. Variable KDE preforms better in finding kernel structures with ISOMAP and LLE, and in locating outliers with t-SNE and UMAP.", echo = FALSE}
knitr::include_graphics("figures/Twin Peak2000_5levels_outliers_comparison_5ml_r0_5.png")
```

@@

```{r tpcors, echo=FALSE, message=FALSE, eval=T}
load("figures/CorrelationTable_Twin Peak2000_5ml_r0_5.rda")
col.max <- apply(cors, 2, max)
max.match <- rep(col.max, each = 2) == cors
data.frame(cors) %>% 
  mutate_if(is.numeric, format, digits = 3, nsmall = 3) %>% 
  kableExtra::kbl(caption = "Correlation between true density ranking and estimated density ranking for different manifold learning embeddings of the twin peak data. Variable bandwidth KDE outperfoms for LLE and UMAP, and LLE gives the highest rank correlation.", booktabs = TRUE, digits = 3, escape = FALSE) %>%
  # kable_styling(latex_options = "scale_down") %>%
  kable_paper(full_width = FALSE) %>%
  # column_spec(1, width = "6cm") %>% 
  column_spec(2, bold = max.match[,1]) %>%
  column_spec(3, bold = max.match[,2]) %>%
  column_spec(4, bold = max.match[,3]) %>%
  column_spec(5, bold = max.match[,4]) %>%
  column_spec(6, bold = max.match[,5])
```

@@

```{r tpisomapden, fig.align = 'center', out.width = "95%", fig.cap="Scatterplot of true density and estimated density ranks of ISOMAP embedding for KDE with both variable and fixed bandwidth. The four colors and shapes represents the four gaussian kernels in the 2-D meta data. DC-KDE shows a strong linear positive relationship.", echo = FALSE}
knitr::include_graphics("figures/Twin Peak2000_density_isomap_r0_5_rank.png")
```

@@

```{r tpisomapvs4ml, fig.align = 'center', out.width = "95%", fig.cap="Comparison of outliers found by ISOMAP and other four manifold learning methods for DC-KDE (on the left panel) and KDE (on the right panel). The four colors and shapes represents the four gaussian kernels in the 2-D meta data. Outliers found by DC-KDE are more consistent regardless of the manifold learning embedding.", echo = FALSE}
knitr::include_graphics("figures/Twin Peak2000_density_comparison_isomapvs4ml_r0_5_rank.png")
```

Finally, \autoref(fig:tpisomapvs4ml) is used to show the robustness of the density estimates using DC-KDE regardless of the embedding method. By plotting the density ranks of LLE, Laplacian Eigenmaps, t-SNE, and UMAP against ISOMAP, we could tell that the rank correlation from DC-KDE is higher that that from KDE, meaning that the rank of DC-KDE are more consistent with the manifold learning methods. Compared with KDE where the density ranks varies with the embedding methods even for the same manifold, DC-KDE is more robust when the distortions are fixed.
